---
title: "Merging Iraq and Syria Shapefiles"
author: "Daniel K Baissa"
date: "3/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

In this document I will merge the Iraq and Syria shapefiles into one joined shapefile.

I will use the following libraries to join the shapefiles:

```{r}
library(rgdal)
library(raster)
```

Let's load in the Iraq Shapefile first. I will use readOGR and then plot the shapefile to make sure it is working.

```{r}
Iraq <- readOGR("Shapefiles/Iraq", "pc284dt8747")
plot(Iraq)
```

It is working. Now let's load in Syria.

```{r}
Syria <- readOGR("Shapefiles/Syria", "gk468fm4741")
plot(Syria)
```

The Syria shapefile also loaded correctly.

### Merging

Now that we have our shapefiles, we can merge them using the bind() function.

```{r}
Iraq_and_Syria <- bind(Iraq, Syria)

plot(Iraq_and_Syria)
```

Now let's save the merged shapefile.

```{r}
writeOGR(Iraq_and_Syria, "Shapefiles/Iraq_Syria_merged", "Iraq_and_Syria", driver = "ESRI Shapefile")
```

### Disolving the Polygons

Now that we have the joined Iraq and Syria shapefile, I want a shapefile that disolves away the borded between the two countries. To do this I will use the aggregate() function.

```{r}
Iraq_Syria_joined <- aggregate(Iraq_and_Syria)

plot(Iraq_Syria_joined)
```

Great! Now we have a polygon with Iraq and Syria joined and the border between the two dissolved.




